---
title: "Data_Manipulation"
author: "Eileen Shea"
date: "9/29/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Import FAS csv files

Import my csv files to use as examples.

```{r}
litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```


## Select variables

...by listing:

```{r}
select(litters_data, group, litter_number, gd0_weight)
select(litters_data, litter_number, gd0_weight, pups_survive)
```

...by specifying a range:

```{r}
select(litters_data, gd_of_birth:pups_survive)
```

...by saying what you want to remove:

```{r}
select(litters_data, -litter_number)
```

...can include renaming: 

```{r}
select(litters_data, group, litter_number, gest_day_0_weight = gd0_weight)
rename(litters_data, gest_day_0_weight = gd0_weight)
```

...use select helpers:

```{r}
select(litters_data, starts_with("gd"))
select(litters_data, litter_number, starts_with("gd"))

select(litters_data, litter_number, everything())
```

Learning assessment: In the pups data, select the columns containing litter number, sex, and PD ears.

```{r}
select(pups_data, litter_number, sex, pd_ears)
```


## Filter observations

Do some filtering!!

```{r}
filter(litters_data, gd0_weight < 25)
filter(litters_data, gd0_weight >= 25)

filter(litters_data, pups_born_alive == 8)

filter(litters_data, !is.na(gd0_weight))

filter(litters_data, group == "Low8" | group == "Low7" )
filter(litters_data, group %in% c("Low7", "Low8"))

filter(litters_data, pups_born_alive %in% 4:6)

filter(litters_data, pups_born_alive %in% 4:7, !is.na(gd0_weight))
```

Learning Assessment: Filter to include only pups with sex 1 ; Filter to include only pups with PD walk less than 11 and sex 2

```{r}
filter(pups_data, sex == 1)

filter(pups_data, pd_walk < 11 & sex == 2)
```


